---
title: "Performance notes"
output: html_document
---

There could be several ways to do stuff. Which one is faster?
See [comparison of single value
assignment](https://github.com/gagin/R-tricks/blob/master/compare-dt-df.R)
in data.frame, data.table and list. It seems that it's faster to read from
higher level construct but write to lower level construct.

```{r, include=FALSE}
library(microbenchmark)
library(ggplot2)
library(magrittr)
library(data.table)
```

## Strsplit is faster than regexp

```{r}
cleanup.strsplit <- function(s) {
        strsplit(x=s, split="oz.) ", fixed=TRUE)[[1]][2]
}

cleanup.sub <- function(s) {
        sub("(.+oz\\.\\) )(.+)", "\\2", s)
}

l <- "(   oz.) tomato paste"
microbenchmark(
        cleanup.sub(l),
        cleanup.strsplit(l),
        times=100L
        ) %>%
        autoplot
```

## row.names vs rownames

In the experiment below,

* rownames are faster for matrix for both read and write,
* it doesn't matter for data.table and data.frame, although row.names is a bit
faster

TODO Check if there are confidence intervals built-in in microbenchmark(),
and implement, if they aren't

```{r, echo=FALSE}
m <- matrix(1:10000, ncol=1, nrow=10000, dimnames=list(1:10000))
m1 <- m
mb.matrix <- microbenchmark(
        row.names(m1) <- row.names(m),
        row.names(m1) <- rownames(m),
        rownames(m1) <- row.names(m),
        rownames(m1) <- rownames(m),
        times = 1000L)     
autoplot(mb.matrix)
mb.matrix

df <- data.frame(m)
df1 <- df
mb.df<-microbenchmark(
        row.names(df1) <- row.names(df),
        row.names(df1) <- rownames(df),
        rownames(df1) <- row.names(df),
        rownames(df1) <- rownames(df),
        times = 1000L)
autoplot(mb.df)
mb.df

dt <- data.table(m)
dt1 <- dt
mb.dt<-microbenchmark(
        row.names(dt1) <- row.names(dt),
        row.names(dt1) <- rownames(dt),
        rownames(dt1) <- row.names(dt),
        rownames(dt1) <- rownames(dt),
        times = 1000L)
autoplot(mb.dt)
mb.dt
```

