---
title: "Performance notes"
output: html_document
---

There could be several ways to do stuff. Which one is faster?
See [comparison of single value
assignment](https://github.com/gagin/R-tricks/blob/master/compare-dt-df.R)
in data.frame, data.table and list. It seems that it's faster to read from
higher level construct but write to lower level construct.

```{r, include=FALSE}
library(microbenchmark)
library(ggplot2)
library(magrittr)
library(data.table)
```

## Strsplit is faster than regexp

```{r}
cleanup.strsplit.NA <- function(s) {
        # This one is broken - it will return NA if the split isn't found
        strsplit(x=s, split="oz.) ", fixed=TRUE)[[1]][2]
}

cleanup.strsplit <- function(s) {
        s1 <- strsplit(x=s, split="oz.) ", fixed=TRUE)[[1]][2]
        if(is.na(s1)) s else s1
}

cleanup.sub <- function(s) {
        sub("(.+oz\\.\\) )(.+)", "\\2", s)
}

l <- rep(c("simple","(   oz.) tomato paste"),10)
microbenchmark(
        lapply(l,cleanup.sub),
        lapply(l,cleanup.strsplit),
        lapply(l,cleanup.strsplit.NA),
        times=1000L
        ) %>%
        autoplot
```

## row.names vs rownames

In the experiment below,

* rownames are faster for matrix for both read and write,
* it doesn't matter for data.table and data.frame, although row.names is a bit
faster

Interesingly, there're row.names and rownames, but there's only colnames.
There's no col.names

TODO Check if there are confidence intervals built-in in microbenchmark(),
and implement, if they aren't

```{r, echo=FALSE}
m <- matrix(1:10000, ncol=1, nrow=10000, dimnames=list(1:10000))
m1 <- m
mb.matrix <- microbenchmark(
        row.names(m1) <- row.names(m),
        row.names(m1) <- rownames(m),
        rownames(m1) <- row.names(m),
        rownames(m1) <- rownames(m),
        times = 1000L)     
autoplot(mb.matrix)
mb.matrix

df <- data.frame(m)
df1 <- df
mb.df<-microbenchmark(
        row.names(df1) <- row.names(df),
        row.names(df1) <- rownames(df),
        rownames(df1) <- row.names(df),
        rownames(df1) <- rownames(df),
        times = 1000L)
autoplot(mb.df)
mb.df

dt <- data.table(m)
dt1 <- dt
mb.dt<-microbenchmark(
        row.names(dt1) <- row.names(dt),
        row.names(dt1) <- rownames(dt),
        rownames(dt1) <- row.names(dt),
        rownames(dt1) <- rownames(dt),
        times = 1000L)
autoplot(mb.dt)
mb.dt
```

## Apply vs for loop

Technically, apply function do the same thing as for loops, but if something
can be organized to be used in apply(), then it's probably more efficient
as well.